What We Eval in the Shadows

### 2021-04-02 paper/main.tex
various changes

\begin{document}
\title{What we {\sf eval} in the shadows}
\subtitle{A large-scale study of {\sf eval} usage in R programs}

@@

%%%%%%%% JAN:::: NOT SURE WHAT THIS IS
% The namespace stuff of packages was meant for the description of the envir argument.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{itemize}
\item \underline{$min(e)=\c{V}$}: Expression that represent values are
  frequently passed to \eval. The majority of those, \packageValOneNodePercent,
  are inline constants (integers or doubles). The rest trivially evaluate to a
  value.\footnote{True as long as base functions such as \c{+} are not
  redefined. They typically aren't, but this is a limitation of this
  categorization.} In our corpus, \packageNbCallSitesUniqueActualValue call
  sites only ever see a simple values; the median number of times they are
  exercised is \packageMedianRunSitesUniqueActualValue, suggesting a possible
  lack of coverage. However, manual inspection reveals cases such as:
\begin{lstlisting}
 f1 <- eval(paste("A~",paste(paste(names(X[,-1])),collapse="+")))
\end{lstlisting}
  This concatenates to a string and \eval simply returns it as is.
  From twenty randomly selected \evals, only \packageUsefulValueEvalPercent of
  them needed to use \eval.
%% The percentage of non default environment passed to \eval is
%% \packageMinimizedpercentenvirc when the minimized expression is a value. Some
%% sites which are passed a value, are also passed different minimized expressions.
%% For those ones, \packageNonDefaultEnvirValuePercent use a non-default
%% environment, and among the one that use a non-default environment,
%% \packageNonDefaultEnvirWithVarPercent have at least one minimized expression
%% with a variable inside.
\item \underline{$min(e)=\c{X}$}: Variables are the most common form. They
  subsume values, thus \c{X} includes arithmetic expressions in which a variable
  occurs. \packageNbSymbolVarSitePercent of variable expression are symbols, \ie
  just an \c{x}. As a variable may refer to a promise, it may trigger
  computation.
  %% The
  %% \c{envir} argument is explicitly passed in \packageMinimizedpercentenvira\% of
  %% the call sites. The call sites without an explicit \c{envir} argument all use
  %% a specific \eval variant, \c{eval.parent}. Moreover, the \c{envir} argument is
  %% both explicit and different from \c{parent.frame()} in
  %% \packageMinimizedpercentenvirnonparenta\% of the sites, which suggests that
  %% when given a variable, \eval is often used to get value in a non-local,
  %% non-parent environment.
\item \underline{$min(e)=\c{\$}$}: A name lookup, \c{x\$f}, can be used to
  access variable in an environment or a column in a data frame, this also
  include vector   indexing operators, \c{[]} and \c{[[]]}. A name lookup
  subsumes variables.
%% It amounts for \packageNbSlotAccess sites, \ie
%% \packageSlotAccessPercent of the sites. The non-default, non-parent environment
%% percentage is here \packageMinimizedpercentenvirnonparentg.
\item \underline{$min(e)=$~\c{<-}}: Assignments account for
  \packageAssignSitesPercent of \eval sites, they are \c{<-}, {\tt <<-} and
  \c{assign}, and they subsume named lookups.
%% \packageNonDefaultEnvirAssignSitesPercent of the sites with assignments. Indeed,
%% most of those assignments occur in the default environment, \ie in the caller of
%% \eval. This hints that side effects are not widespread in \eval.\footnote{Side
%% effects can also happen in a function call in the \eval expression, and that
%% function call will hide assignment operators from us. We more precisely
%% instrument side-effects to environment in Section~REF.}
\item \underline{$min(e)=\c{F()}$}: A plain function call which may have some
  values as argument but does not include variables. It should be noted that
  function name require variable lookups -- we distinguish those as they usually
  do not trigger computation.
\item \underline{$min(e)=\c{F(X)}$}: Function calls whose arguments may include
  variable references.
\item \underline{$min(e)=\c{F(F(X))}$}: Nested function calls with variables.
\item \underline{$min(e)=\c{FUN}$}: Function definition accounts for
  \packageFunctionDefinitionSitesPercent of \eval sites. In addition the
  functions inside other expressions account for
  \packageGeneralizedFunctionDefinitionSitesPercent of sites, and give an
  estimate of the use of higher-order functions.
\item \underline{$min(e)=\c{\{\,\}}$}: Arbitrary-length statement blocks. These
  are larger expressions with, on average, \packageMinimizedoperationsjRnd
  instructions. The contents of the block are not inspected.
\item \underline{$min(e)=\c{model.frame}$}: Call to the \c{model.frame} function
  used to create statistical models. This subsume all other categories.

@@

%% Similarly, when a function is evaluated on some variable (\c{F(X)} minimized
%% expression), the percentage of non default, non parent environments is rather
%% high, \packageMinimizedpercentenvirnonparentd\%.

%% As a whole, variable access, either as the top call, or inside a function call,
%% possibly with accessing a slot, represents
%% \packageGeneralizedVarAccessSitePercent of the sites.

%% The minimized expression classification suggests that the presence of variables
%% in the expression is highly correlated to the specification of a specific
%% environment, and that \eval is used here for its power in precising the
%% evaluation environment.

@@

The \c{envir} argument of \eval can be an actual environment, but also \c{Null},
a list, a dataframe, a pairlist or even an integer. Most of the time, \ie in
\packageEnvirEnvSitePercent of the sites, \eval is passed an actual environment.
In \packageNbEnvirIncorrectSites sites (\packageEnvirIncorrectSitePercent),
\eval is passed an incorrect argument, for instance \c{NA}, a boolean, which
makes \eval fail. \c{NULL}, a \c{list} or a data frame, or an integer, happen in
\packageEnvirUnusualSitePercent of the cases: \eval copies the fields of the
list or data frame and creates bindings for them in a new environment. This
pattern is used to evaluate formulas which can directly refer to the fields of
the data. The \c{envir} argument can also be a number $n$. It means that the
environment in which the expressions is evaluated will be the result of
\c{sys.call(n)} where $n$ refers to the $n$-th stack frame.

@@

@@

The \c{imchange} function of package \c{imager} makes it possible to modify
images using a dedicated formula syntax using \~.\footnote{Inspired by {map} in
package \emph{purr}.} Here, \eval is evaluated in \c{newenv}, which creates a
new environment that inherits from \c{parent.frame()} by default (classified as
1+).

@@

\paragraph{Package environments.} Another kind of environments that is passed to \eval is the environment of a package.
 It refers to the \emph{namespace} of a package. A call to
 \c{library(packageName)}, which loads libraries in R, for instance, will
 populate environment \c{package:packageName}. It is a \emph{base} function
 though, so we do not see it in the CRAN dataset. However, R still makes it
 possible to access any package environments and \packageNbPackageEnvSites call
 sites, \ie \packagePackageEnvSitePercent access a package environment, from
 \packageNbPackageEnvPackages different packages. Those packages perform
 introspection, to represent and search any environment by name, in package
 \emph{envnames}, to containerize tests, as with \emph{testthat} and
 \emph{unitizer}, to create a custom OOP system with \emph{R.oo}, or to create
 poor man's hidden functions in \emph{USSCXenaTools}.

 It is also used to implement a custom caching mechanism in packages
 \emph{g.data} and \emph{SOAR}.

\paragraph{\c{enclos} argument.}

\mypara{Side-effects.} XXX026 total number of reads performed by
\eval. \NOTE{What is the definition of a read? An env lookup?} XXX027
average number of reads. XXX028 total number of writes performed by
\eval. Average number XXX029.

\NOTE{Aviral wrote''measure number of reads and writes by eval outside
  of its bubble'' --- what is the bubble here?}

XXX029 reads to caller environment -- where caller is the caller of
\eval.

XXX030 reads to other caller environment -- i.e. envs that are on the
call stack.

XXX031 reads to caller environment -- where caller is the caller of
\eval.

XXX032 reads to other caller environment -- i.e. envs that are on the
call stack.

@@

\subsection{Comparison with Javascript}

\begin{table}[h]
	\begin{tabularx}{\textwidth}{|X|X|X|X|X|}
		\hline
		& Julia & Javascript & Python & R \tabularnewline
		\hline % lstlinline does not like tabular environments...
		{\eval} & \texttt{eval(expr)} & \texttt{eval(expr)} & \texttt{eval(expr[, globals[, locals]])} &  \texttt{eval(expr[, envir[, enclos]])}  \tabularnewline
		\hline
		Environments &Global  & Local & Global or local  & Any \tabularnewline
		\hline
	\end{tabularx}
	\caption{Relative power of \eval throughout languages: in which environment it can evaluate its expression.}
\end{table}

>>>>

\begin{document}
\title{What We Eval in the Shadows}
\subtitle{A large-scale study of {\sf eval} usage in R programs}

@@

@@

\begin{itemize}
\item \underline{$min(e)=\c{V}$}: Expression that represent values are
  frequently passed to \eval. The majority of those, \packageValOneNodePercent,
  are inline constants (integers or doubles). The rest trivially evaluate to a
  value.\footnote{True as long as base functions such as \c{+} are not
  redefined. They typically aren't, but this is a limitation of this
  categorization.} In our corpus, \packageNbCallSitesUniqueActualValue call
  sites only ever see a simple values; the median number of times they are
  exercised is \packageMedianRunSitesUniqueActualValue, suggesting a possible
  lack of coverage. However, manual inspection reveals cases such as:
\begin{lstlisting}
 f1 <- eval(paste("A~",paste(paste(names(X[,-1])),collapse="+")))
\end{lstlisting}
  This concatenates to a string and \eval simply returns it as is.
  From twenty randomly selected \evals, only \packageUsefulValueEvalPercent of
  them needed to use \eval.
\item \underline{$min(e)=\c{X}$}: Variables are the most common form. They
  subsume values, thus \c{X} includes arithmetic expressions in which a variable
  occurs. \packageNbSymbolVarSitePercent of variable expression are symbols, \ie
  just an \c{x}. As a variable may refer to a promise, it may trigger
  computation.
\item \underline{$min(e)=\c{\$}$}: A name lookup, \c{x\$f}, can be used to
  access variable in an environment or a column in a data frame, this also
  include vector   indexing operators, \c{[]} and \c{[[]]}. A name lookup
  subsumes variables.
\item \underline{$min(e)=$~\c{<-}}: Assignments account for
  \packageAssignSitesPercent of \eval sites, they are \c{<-}, {\tt <<-} and
  \c{assign}, and they subsume named lookups.
\item \underline{$min(e)=\c{F()}$}: A plain function call which may have some
  values as argument but does not include variables. It should be noted that
  function name require variable lookups -- we distinguish those as they usually
  do not trigger computation.
\item \underline{$min(e)=\c{F(X)}$}: Function calls whose arguments may include
  variable references.
\item \underline{$min(e)=\c{F(F(X))}$}: Nested function calls with variables.
\item \underline{$min(e)=\c{FUN}$}: Function definition accounts for
  \packageFunctionDefinitionSitesPercent of \eval sites. In addition the
  functions inside other expressions account for
  \packageGeneralizedFunctionDefinitionSitesPercent of sites, and give an
  estimate of the use of higher-order functions.
\item \underline{$min(e)=\c{\{\,\}}$}: Arbitrary-length statement blocks. These
  are larger expressions with, on average, \packageMinimizedoperationsjRnd
  instructions. The contents of the block are not inspected.
\item \underline{$min(e)=\c{model.frame}$}: Call to the \c{model.frame} function
  used to create statistical models. This subsume all other categories.

@@

@@

The environment in which evaluation happens determines what will be visible to
the computation started by \eval, and the potential reach of side-effects. The
second argument to \eval, \c{envir}, defines the evaluation environment of the
expression. That argument can be omitted, in which case the default value is
\c{parent.frame()} which returns the caller of the \eval function.

Other possible values for the environment include, lists, dataframes, pairlists,
integers and \c{Null}. Lists and dataframe are coerced to new environments,
integers values are passed to the \c{sys.call} function which returns the
environment of a stack frame at the requested offset.

@@

Another kind of environments that is passed to \eval is the environment of a
package. It refers to the \emph{namespace} of a package. A call to
\c{library(packageName)}, which loads libraries in R, for instance, will
populate environment \c{package:packageName}. It is a \emph{base} function
though, so we do not see it in the CRAN dataset. However, R still makes it
possible to access any package environments and \packageNbPackageEnvSites call
sites, \ie \packagePackageEnvSitePercent access a package environment, from
\packageNbPackageEnvPackages different packages. Those packages perform
introspection, to represent and search any environment by name, in package
\emph{envnames}, to containerize tests, as with \emph{testthat} and
\emph{unitizer}, to create a custom OOP system with \emph{R.oo}, or to create
poor man's hidden functions in \emph{USSCXenaTools}. It is also used to
implement a custom caching mechanism in packages \emph{g.data} and \emph{SOAR}.

\subsubsection{Bits of text, possibly reusable}

Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

In \packageNbEnvirIncorrectSites sites (\packageEnvirIncorrectSitePercent),
\eval is passed an incorrect argument, for instance \c{NA}, a boolean, which
makes \eval fail.

\c{NULL}, a \c{list} or a data frame, or an integer, happen in
\packageEnvirUnusualSitePercent of the cases: \eval copies the fields of the
list or data frame and creates bindings for them in a new environment.

This pattern is used to evaluate formulas which can directly refer to the fields
of the data.

@@

The \c{imchange} function of package \c{imager} makes it possible to modify
images using a dedicated formula syntax.\footnote{Inspired by {map} in package
\emph{purr}.} Here, \eval is evaluated in \c{newenv}, which creates a new
environment that inherits from \c{parent.frame()} by default (classified as 1+).

@@

\paragraph{\c{enclos} argument.} \ \\

@@
